The role of viscosity on drop impact forces

### 2024-06-08 Rebuttal/Response to Reviewer 1.tex

%\documentclass[final]{article}
\documentclass[]{article}
\usepackage{geometry}
\geometry{
	a4paper,
	total={170mm,257mm},
	left=0.75in,
	top=0.75in,
	right=0.75in,
	bottom=1in,
}
\newcommand{\extramargin}[1]{}
%\newcommand{\extramargin}[1]{
%	\setlength{\paperwidth}{250mm}
%	\setlength{\evensidemargin}{-15mm}
%	\setlength{\oddsidemargin}{-15mm}
%}
\usepackage{gensymb}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{epstopdf, epsfig}
\usepackage{amsmath}
\usepackage[url=false,
backend=bibtex,
style=authoryear-comp,
giveninits=true,
doi=true,
isbn=true,
backref=false,
dashed=false,
maxcitenames=2,
maxbibnames=99,
natbib=true]{biblatex}
\DeclareNameAlias{author}{last-first}
\DeclareFieldFormat
[article,inbook,incollection,inproceedings,patent,thesis,
unpublished,techreport,misc,book]
{title}{#1}
\renewcommand*{\revsdnamepunct}{}
\renewbibmacro{in:}{}
\usepackage{xpatch}
\xpatchbibmacro{date+extradate}{%
	\printtext[parens]%
}{%
	\setunit*{\space}%
	\printtext%
}{}{}

\addbibresource{../Main/ViscousDropImpact_v2.bib}
\nonfrenchspacing
\setlength\parindent{0pt}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{mathrsfs}
\usepackage[colorlinks,citecolor=purple]{hyperref}
\newcommand*\blue{\textcolor{blue}}
\newcommand*\red{\textcolor{red}}
\newcommand{\VS}[1]{{\textcolor{orange}{#1}}}

\renewcommand{\thefigure}{R\arabic{figure}}

\usepackage[obeyFinal,colorinlistoftodos, textwidth=60mm, shadow]{todonotes}
% todonotes specific macros begin!
\newcommand{\todoInsertref}[2]{\todo[color=green!40, #1]{#2}}
\newcommand{\todoExplaininDetails}[2]{\todo[color=orange!40, #1]{#2}}
\newcommand{\todoUrgent}[2]{\todo[color=red!40, #1]{#2}}
\newcommand{\todoSansUrgent}[2]{\todo[color=yellow!40, #1]{#2}}
\extramargin
%% todonotes specific macros end!

%opening
\title{Reply to Referee 1}
\author{Vatsal Sanjay, Bin Zhang, Cunjing Lv, and Detlef Lohse}
\date{}
\begin{document}
	\maketitle

\textit{The authors describe a combined numerical and experimental study on the influence of Weber and Ohnesorge number on the peak forces (and their associated timing) associated with droplet impact and rebound on a non-wetting, rigid, planar surface. Scaling arguments are proposed throughout to help rationalize the trends observed in the data. This work represents a direct extension of work previously published by the same group (Zhang et al. PRL 2022) wherein they identified the possibility of two peaks in the impact force evolution associated with the initial impact and retraction stages. Overall, the manuscript is well written and easy to follow. However I do have some major and minor concerns that are described in what follows.}\\[2mm]

We thank the referee for carefully reading our manuscript and providing valuable feedback and suggestions. We have reviewed the referee's comments and made changes based on their suggestions. Below, we offer a point-to-point reply to each of the referee's comments and include the changes made in the manuscript. The referee's comments are in italics, and our replies are in plain black. Changes in the manuscript are highlighted in orange.

\begin{enumerate}
	\item \textit{Apart from Figure 2 (which appears to be the previously published data from Figure 1 in Zhang et al. (PRL 2022)), there is very little direct comparison between experiment and simulation. Now that the parameter space has been significantly extended from prior work, I would have expected similar direct force versus time comparisons and droplet shape profiles spanning the new range of parameters to validate the simulation. Comparisons similar to that in Figure 2 should be included throughout when the various parameter various regimes are introduced and discussed.}\\[0.5mm]

	Yes, it is correct that the earlier version of figure 2 summarized the results previously published in \citet{zhang2022impact}. To extend the comparison at different Ohnesorge numbers $Oh$, we have added two additional $F(t)$ curves in figure 2 (see figure~\ref{fig:summary}). Furthermore, throughout the manuscript, we present experimental data wherever available (see figures 3 and 7).

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{../Main/Figures/Figure1_summary_v5.pdf}
		\caption{\VS{Comparison of the drop impact force $F(t)$ obtained from experiments and simulations for the three typical cases with impact velocity $V_0 = 1.2\,\si{\meter}/\si{\second}, 0.97\,\si{\meter}/\si{\second}, 0.96\,\si{\meter}/\si{\second}$, diameter $D_0 = 2.05\,\si{\milli\meter}, 2.52\,\si{\milli\meter}, 2.54\,\si{\milli\meter}$, surface tension $\gamma = 72\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}$ and viscosity $\eta_d = 1\,\si{\milli\pascal\second}, 25.3\,\si{\milli\pascal\second}, 80.2\,\si{\milli\pascal\second}$. These parameter give $Oh = 0.0025, 0.0625, 0.2$ and $We = 40$.
			For the three cases, the two peak amplitudes, $F_1/\rho_dV_0^2D_0^2 \approx$ 0.82, 0.92, 0.99 at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$ and $F_2/\rho_dV_0^2D_0^2 \approx$ 0.37, 0.337, 0.1 at $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$, characterize the inertial shock from impact and the Worthington jet before takeoff, respectively. 
			The drop reaches the maximum spreading at $t_{\text{max}}$ when it momentarily stops and retracts until $t_3 \approx 0.8\sqrt{\rho_dD_0^3/\gamma}$ when the drop takes off ($F = 0$).
			Six instances are further elaborated through numerical simulations for ($We = 40, Oh = 0.0025$), namely (i) $t = 0\,\si{\milli\second}$ (touch-down), (ii) $t = 0.37\,\si{\milli\second}$ ($t_1$), (iii) $t = 2.5\,\si{\milli\second}$ ($t_{\text{max}}$), (iv) $t = 3.93\,\si{\milli\second}$ ($\approx 0.85t_2$), (v) $t = 4.63\,\si{\milli\second}$ ($t_2$), and (vi) $t = 5.25\,\si{\milli\second}$ ($\approx 1.15t_2$). 
			We stress the excellent agreement between experiments and simulations without any free parameters. The insets show representative snapshots at specific time instants overlaid with the drop boundaries from simulations in orange, also revealing good agreement. 
			The left part of each numerical snapshot shows the dimensionless local viscous dissipation function $\tilde{\xi}_\eta \equiv \xi_\eta D_0/\left(\rho_dV_0^3\right) = 2Oh\left(\boldsymbol{\tilde{\mathcal{D}}:\tilde{\mathcal{D}}}\right)$, where, $\boldsymbol{\mathcal{D}}$ is the symmetric part of the velocity gradient tensor, on a $\log_{10}$ scale and the right part the velocity field magnitude normalized with the impact velocity. The black velocity vectors are plotted in the center of mass reference frame of the drop to clearly elucidate the internal flow.}}
		\label{fig:summary}
	\end{figure}

	For the high Ohnesorge number regime ($Oh \gg 1$) discussed in Figure 5, we, unfortunately, do not have corresponding experimental data as we used water-glycerol mixtures in the experiments, which can at most be 1000 times (100\% by weight of glycerine) more viscous than water. However, to provide an additional comparison, we have added a new panel (Figure 5c, see figure~\ref{fig:F1Anatomy_2}) comparing our results with those from \citet{Gordillo2018, cheng2021drop}. This comparison focuses on the relationship between the first peak force $F_1$ and the impact Reynolds number $Re = \rho_dV_0D_0/\eta_d$.

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{../main/Figures/ForceTimeVersusOhnesorge.pdf}
		\caption{\blue{Anatomy of the first impact force peak amplitude for viscous impacts from our numerical simulations: the $Oh$ dependence of  (a) the magnitude $F_1$ normalized by the inertial force scale $\rho_dV_0^2D_0^2$ and (b) the time $t_1$ to reach the first force peak amplitude normalized by inertial timescale $\tau_\rho = D_0/V_0$}. \VS{(c) The $Re$ dependence of the magnitude $F_1$ normalized by the inertial force scale $\rho_dV_0^2D_0^2$ as compared to the (implicit) theoretical calculation of \citet{Gordillo2018}.} \blue{The black line corresponds to the scaling relationship described in \S~3.2. The Weber number is color-coded.}}
		\label{fig:F1Anatomy_2}
	\end{figure}

	\item \textit{Error bars (both horizontal and vertical) for the experiments should be included throughout whenever experimental data is presented. If smaller than the data points, this should be noted. Furthermore, a discussion of any error analysis is completely lacking. For the error bars that are included, do these account for multiple trials? Parameter uncertainties? Measurement uncertainties? Such a discussion, with errors quantified (and methods used to assess the errors), must be included. In particular, the resolution of the force sensor (reported as 0.5 mN) should be indicated when force measurements are presented, including in Figure 2.}\\[0.5mm]

	Thank you for highlighting the importance of error analysis and the need for a more comprehensive discussion of error quantification in our manuscript. We appreciate your suggestions. Indeed, we have included the error bars wherever experimental data are reported. For the cases where the error bars are not visible, the error bar is smaller than the data points. For more details regarding the error quantification, we refer the readers to the supplementary material of \citet{zhang2022impact}. We have added this clarification in the experimental method section:

	\S~\red{2.1:}\\
	\VS{Throughout the manuscript, the error bars account for repeated trials and are visible if they are larger than the marker size. We refer the readers to the supplementary material of \citet{zhang2022impact} for further details of the experimental setup.}\\[0.5mm]

	We have also added a dashed gray horizontal line in figure 2 to mark the resolution of the force sensor.

	\item \textit{Furthermore, the dimensional parameters of the experiments performed need to somehow be specified or indicated. The authors mention a range of droplet sizes, viscosities, and impact speeds, and so We and Oh do not uniquely defined the experimental realizations reported. This specificity is critical for repeatability and future comparisons that might be made.}\\[0.5mm]

	Thank you for raising this important point about the need to specify the dimensional parameters of the experiments. We appreciate the emphasis on the importance of this information for repeatability and future comparisons. We vary the impact velocity by changing the release height of the drops above the substrate and change the viscosity by using different water-glycerol mixtures. Such a protocol allows us to vary the viscosity by almost two orders of magnitude, yet maintaining a fairly constant density and surface tension. Consequently, the Weber number $We$ can be interpreted as a dimensionless impact velocity and the Ohnesorge number $Oh$ as a dimensionless viscosity. We have modified the text to clarify the experimental protocol and provide more details on the dimensional parameters:

	\S~\red{2.1:}\\
	\VS{The properties of the impacting drop are controlled using water-glycerol mixtures with viscosities $\eta_d$ varying by almost two orders of magnitude, from $1\,\si{\milli\pascal}\si{\second}$ to $80.2\,\si{\milli\pascal}\si{\second}$, yet maintaining a fairly constant surface tension $\gamma$ and density $\rho_d$ around $61\,\si{\milli\newton}/\si{\meter}$ and $1000\,\si{\kilo\gram}/\si{\cubic\meter}$, respectively \citep{cheng2008formula, volk2018density, Jha2020}. Using other liquids like silicone oil could allow for a wider viscosity variation if used with a superamphiphobic substrate \citep{deng2012candle}. The drop diameter $D_0$ is controlled between $2.05\,\si{\milli\meter}$ and $2.76\,\si{\milli\meter}$ by pushing it through a calibrated needle. Throughout this work, we use drops with diameter $2.54 \pm 0.02\,\si{\milli\meter}$ unless otherwise stated. Consequently, changing the viscosity of the drop determines the Ohnesorge number ($Oh$, see~1.2). The Weber number ($We$, see~1.1) is set using the impact velocity $V_0$ varying between $0.38\,\si{\meter}/\si{\second}$ and $2.96\,\si{\meter}/\si{\second}$ by changing the release height of the drops above the substrate. Lastly, we keep the Bond number ($Bo$, see~1.3) fixed at 1 throughout the manuscript. In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/gD_0 = We/Bo$ comparing inertia with gravity. Throughout this manuscript, $Fr > 1$ and gravity's role is sub-dominant compared to inertia.}

	This expanded description provides several key pieces of information:
	\begin{enumerate}
		\item The range of viscosities ($1,\si{\milli\pascal}\si{\second}$ to $80.2,\si{\milli\pascal}\si{\second}$) achieved using water-glycerol mixtures.
		\item The relatively constant surface tension ($61,\si{\milli\newton}/\si{\meter}$) and density ($1000,\si{\kilo\gram}/\si{\cubic\meter}$) of these mixtures.
		\item The range of drop diameters ($2.05,\si{\milli\meter}$ to $2.76,\si{\milli\meter}$) and the typical diameter used ($2.54 \pm 0.02,\si{\milli\meter}$).
		\item The range of impact velocities ($0.38,\si{\meter}/\si{\second}$ to $2.96,\si{\meter}/\si{\second}$) achieved by varying the release height.
	\end{enumerate}

	We also clarify that the Ohnesorge number ($Oh$) is uniquely determined by the drop viscosity, while the Weber number ($We$) is set by the impact velocity.

	By providing these details, we aim to ensure that our experiments can be readily reproduced and that future comparisons can be made with a clear understanding of the dimensional parameters involved. We believe this additional information significantly enhances the repeatability and usefulness of our work.

	We have also modified the caption of figure 2 (see figure~\ref{fig:summary}) to clarify the drop impact parameters.

	\item \textit{Is gravity included in the simulations? If so, the Bond number in simulations should be specified, as We and Oh alone do not uniquely define the non-dimensional problem. The authors show that droplet rebound can be affected by finite Bond number effects in their very recent prior work (Sanjay et al. JFM 2023), so at least some discussion on the role of gravity should be included.}\\[0.5mm]

	Thank you for raising this important question about the role of gravity in our simulations and the need to specify the Bond number. In deed, in \citet{sanjay_chantelot_lohse_2023}, we highlight the potential influence of finite Bond number on the bouncing--no-bouncing transition on a non-wetting surface.

	To address your concerns, we have made the following revisions to the manuscript:

	\begin{enumerate}
		\item We have clarified that gravity is indeed included in our simulations and that the Bond number (based on the drop's diameter) $Bo = \rho_dgD_0^2/\gamma$ is fixed at 1 throughout the study. This information has been added to the manuscript as follows:

		\S~\red{2.1:}\\
		\VS{Throughout the manuscript, we keep the Bond number ($Bo$, see~(1.3)) fixed at 1 in our simulations, which include the effects of gravity.}

		\item We have updated Figure 10 to indicate that the critical $Oh$ to suppress bouncing is 0.53 for the chosen $Bo$ of 1. Note that in \citet{sanjay_chantelot_lohse_2023}, we use the radius $R_0 = D_0/2$ as the length scale to get $Oh_c + Bo_c \approx 1$.

		\item We have introduced the dimensionless Froude number $Fr = V_0^2/gD_0 = We/Bo$, which compares inertia with gravity. We clarify that throughout the manuscript, $Fr > 1$, indicating that the role of gravity is sub-dominant compared to inertia. This discussion has been added as follows:

		\S~\red{2.1:}\\
		\VS{In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/gD_0 = We/Bo$, which compares inertia with gravity. Throughout this manuscript, $Fr > 1$, indicating that the role of gravity is sub-dominant compared to inertia.}
	\end{enumerate}

	\item \textit{The title needs to be rewritten. The authors demonstrate that capillarity also plays a crucial role, and seem to ignore gravity throughout the present work anyways (i.e. gravity doesn’t dictate impact forces here, because it is assumed). Furthermore, the title does not specify the nature of the substrate, which is a specific case. I would suggest something along the lines of ``Impact forces of droplets on non-wetting planar surfaces" or similar.}

	We agree with the reviewer that the title was misleading. Of course, all three: inertia, capillarity, and viscosity, play a role. Indeed, we have studied the role of inertial and capillarity in \citet{zhang2022impact}. We have now changed the title of the paper to:

	\VS{The role of viscosity on drop impact forces}

	\item \textit{What are the receding and advancing contact angles for the experimental surfaces used? The experimental substrate is not truly non-wetting (in contrast to the simulation), so this should be clearly described.}\\[0.5mm]

	Yes, the reviewer is correct. The substrate is not truly non-wetting in the experiments. We have added this in the revised manuscript along with the advancing and the receding contact angles.

	\S~\red{2.1:}\\
	 \VS{For water drops, such a surface is coated with silanized silica nanobeads with a diameter of $20\,\si{\nano\meter}$ (Glaco Mirror Coat Zero; Soft99) resulting in the advancing and receding contact angles of $167 \pm 2^{\circ}$ and $154 \pm 2^{\circ}$, respectively \citep{Gauthier2015, Li2017}. On the other hand, for viscous aqueous glycerin drops, the upper surface is coated with an acetone solution of hydrophobic beads (Ultra ever Dry, Ultratech International, a typical bead size of 20 nm), resulting in the advancing and receding contact angles of $166 \pm 4^{\circ}$ and $159 \pm 2^{\circ}$, respectively \citep{Jha2020}.}

	\item \textit{In the experiments, are the droplets arriving at the surface without residual oscillation? Particular for the small We, small Oh cases, I would expect this is not the case. This should be discussed.}

	Thank you for raising this important point about the potential influence of residual oscillations in the drop as they arrive at the surface, particularly for cases with small Weber and Ohnesorge numbers. We agree that this aspect of the experimental setup deserves discussion in the manuscript.
	To address this concern, we have added the following text to the manuscript:

### 2024-06-17 Main/ViscousDropImpact_v7.tex
started editing

%\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\VS}[1]{{\textcolor{orange}{#1}}}

%\shorttitle{Inertia \& viscosity dictate drop impact forces}
\shorttitle{\VS{The role of viscosity on drop impact forces}}
\shortauthor{V. Sanjay, B. Zhang, C. Lv, and D. Lohse}

%\title{Inertia \& viscosity dictate drop impact forces}
\title{\VS{The role of viscosity on drop impact forces}}
\author{Vatsal Sanjay\aff{1}
 \corresp{\email{vatsalsanjay@gmail.com}},
 Bin Zhang\aff{2}
 \corresp{\email{binzhang0710@mail.tsinghua.edu.cn}},
 Cunjing Lv\aff{2}
 \corresp{\email{cunjinglv@mail.tsinghua.edu.cn}},
 \and Detlef Lohse{\aff{1}$^{,}$\aff{3}}
 \corresp{\email{d.lohse@utwente.nl}}}

@@

A liquid drop impacting a rigid substrate undergoes deformation and spreading due to normal reaction forces, which are counteracted by surface tension. On a non-wetting substrate, the drop subsequently retracts and takes off.
Our recent work (Zhang et al., \textit{Phys. Rev. Lett.}, vol. 129, 2022, 104501) revealed two peaks in the temporal evolution of the normal force $F(t)$--one at impact and another at jump-off. The second peak coincides with a Worthington jet formation, which vanishes at high viscosities due to increased viscous dissipation affecting flow focusing.
In this article, using experiments, direct numerical simulations, and scaling arguments, we characterize both the peak amplitude $F_1$ at impact and the one at take off ($F_2$) and elucidate their dependency on the control parameters: the Weber number $We$ (dimensionless impact velocity) and the Ohnesorge number $Oh$ (dimensionless viscosity). 
\VS{The first peak amplitude $F_1$ and the time $t_1$ to reach it depend on inertial timescales for low viscosity liquids, remaining nearly constant for viscosities up to 100 times that of water. For high viscosity liquids, we balance the rate of change in kinetic energy with viscous dissipation to obtain new scaling laws: $F_1/F_\rho \sim \sqrt{Oh}$ and $t_1/\tau_\rho \sim 1/\sqrt{Oh}$, where $F_\rho$ and $\tau_\rho$ are the inertial force and time scales, respectively, which are consistent with our data.}
The time $t_2$ at which the amplitude $F_2$ appears is set by the inertio-capillary timescale $\tau_\gamma$, independent of both the viscosity and the impact velocity of the drop. However, these properties dictate the magnitude of this amplitude.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Schematic_v2.pdf}
	\caption{(a) Problem schematic with an axisymmetric computational domain used to study the impact of a drop with diameter $D_0$ and velocity $V_0$ on a non-wetting substrate. In the experiments, we use a quartz force sensor to measure the temporal variation of the impact force. The subscripts $d$ and $a$ denote the drop and air, respectively, to distinguish their material properties, which are the density $\rho$ and the dynamic viscosity $\eta$. The drop–air surface tension coefficient is $\gamma$. The grey dashed-dotted line represents the axis of symmetry, $r = 0$. Boundary air outflow is applied at the top and side boundaries (tangential stresses, normal velocity gradient, and ambient pressure are set to zero). The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). (b) The phase space with control parameters: the Weber number ($We$: dimensionless impact velocity) and the Ohnesorge number ($Oh$: dimensionless viscosity), exemplifying different applications.}
	\label{fig:schematic}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Figure1_summary_v5.pdf}
	\caption{\VS{Comparison of the drop impact force $F(t)$ obtained from experiments and simulations for the three typical cases with impact velocity $V_0 = 1.2\,\si{\meter}/\si{\second}, 0.97\,\si{\meter}/\si{\second}, 0.96\,\si{\meter}/\si{\second}$, diameter $D_0 = 2.05\,\si{\milli\meter}, 2.52\,\si{\milli\meter}, 2.54\,\si{\milli\meter}$, surface tension $\gamma = 72\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}$ and viscosity $\eta_d = 1\,\si{\milli\pascal\second}, 25.3\,\si{\milli\pascal\second}, 80.2\,\si{\milli\pascal\second}$. These parameter give $Oh = 0.0025, 0.0625, 0.2$ and $We = 40$.
	For the three cases, the two peak amplitudes, $F_1/\rho_dV_0^2D_0^2 \approx$ 0.82, 0.92, 0.99 at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$ and $F_2/\rho_dV_0^2D_0^2 \approx$ 0.37, 0.337, 0.1 at $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$, characterize the inertial shock from impact and the Worthington jet before takeoff, respectively. 
	The drop reaches the maximum spreading at $t_{\text{max}}$ when it momentarily stops and retracts until $t_3 \approx 0.8\sqrt{\rho_dD_0^3/\gamma}$ when the drop takes off ($F = 0$).
	Six instances are further elaborated through numerical simulations for ($We = 40, Oh = 0.0025$), namely (i) $t = 0\,\si{\milli\second}$ (touch-down), (ii) $t = 0.37\,\si{\milli\second}$ ($t_1$), (iii) $t = 2.5\,\si{\milli\second}$ ($t_{\text{max}}$), (iv) $t = 3.93\,\si{\milli\second}$ ($\approx 0.85t_2$), (v) $t = 4.63\,\si{\milli\second}$ ($t_2$), and (vi) $t = 5.25\,\si{\milli\second}$ ($\approx 1.15t_2$). 
	We stress the excellent agreement between experiments and simulations without any free parameters. The insets show representative snapshots at specific time instants overlaid with the drop boundaries from simulations in orange, also revealing good agreement. 
	The left part of each numerical snapshot shows the dimensionless local viscous dissipation function $\tilde{\xi}_\eta \equiv \xi_\eta D_0/\left(\rho_dV_0^3\right) = 2Oh\left(\boldsymbol{\tilde{\mathcal{D}}:\tilde{\mathcal{D}}}\right)$, where, $\boldsymbol{\mathcal{D}}$ is the symmetric part of the velocity gradient tensor, on a $\log_{10}$ scale and the right part the velocity field magnitude normalized with the impact velocity. The black velocity vectors are plotted in the center of mass reference frame of the drop to clearly elucidate the internal flow.}}
\label{fig:summary}
\end{figure}

@@

Naturally, even the process of a Newtonian liquid drop impacting a rigid substrate is governed by a plethora of control parameters, including but not limited to the drop's density $\rho_d$, diameter $D_0$, velocity $V_0$, dynamic viscosity $\eta_d$, surface tension $\gamma$, \VS{and acceleration due to gravity $g$} (figure~\ref{fig:schematic}a). 
To navigate this rich landscape, we focus on two main dimensionless numbers that serve as control parameters (figure~\ref{fig:schematic}b): the Weber number $We$, which is the ratio of inertial to capillary forces and is given by

@@

\noindent \VS{Additionally, the Bond number }

\VS{\begin{align}
		\label{Eq:BoDefinition}
	Bo = \frac{\rho_dgD_0^2}{\gamma}
\end{align}}

\noindent \VS{compares gravity to inertial forces and is needed to uniquely define the non-dimensional problem.}

@@

\begin{table}
	\begin{center}
		\def~{\hphantom{0}}
		\begin{tabular}{lccc}
			glycerol & $\rho_d$ & $\eta_d$  & $\gamma$\\
			(wt \%) &(kg/m$^{3}$) &(mPa.s)& (mN/m)\\[3pt]
			0 & 1000 & 1 & 72 \\
			50 & 1124 & 5 & 61 \\
			63 & 1158 & 10 & 61 \\
			74 & 1188 & 25.3 & 61 \\
			80 & 1200 & 45.4 & 61 \\
			85 & 1220 & 80.2 & 61 \\
		\end{tabular}
		\caption{\VS{Properties of the water-glycerol mixtures used in the experiments. $\rho_d$ and $\eta_d$ are the density and viscosity of the drop, respectively and $\gamma$ denotes the liquid-air surface tension coefficient. These properties are calculated using the protocol provided in \citet{cheng2008formula, volk2018density}.}}
		\label{tab:table00}
	\end{center}
\end{table}

>>>>

%\usepackage[normalem]{ulem}
\usepackage{xcolor}
\newcommand{\VS}[1]{{\textcolor{orange}{#1}}}
\newcommand{\red}[1]{{\textcolor{red}{#1}}}

%\shorttitle{Inertia \& viscosity dictate drop impact forces}
\shorttitle{The role of viscosity on drop impact forces}
\shortauthor{V. Sanjay, B. Zhang, C. Lv, and D. Lohse}

%\title{Inertia \& viscosity dictate drop impact forces}
\title{The role of viscosity on drop impact forces}
\author{Vatsal Sanjay\aff{1}
 \corresp{\email{vatsalsanjay@gmail.com}},
 Bin Zhang\aff{2}
 \corresp{\email{binzhang0710@mail.tsinghua.edu.cn}},
 Cunjing Lv\aff{2}
 \corresp{\email{cunjinglv@mail.tsinghua.edu.cn}},
 \and Detlef Lohse{\aff{1}$^{,}$\aff{3}}
 \corresp{\email{d.lohse@utwente.nl}}}

@@

A liquid drop impacting a rigid substrate undergoes deformation and spreading due to normal reaction forces, which are counteracted by surface tension. On a non-wetting substrate, the drop subsequently retracts and takes off.
Our recent work (Zhang et al., \textit{Phys. Rev. Lett.}, vol. 129, 2022, 104501) revealed two peaks in the temporal evolution of the normal force $F(t)$--one at impact and another at jump-off. The second peak coincides with a Worthington jet formation, which vanishes at high viscosities due to increased viscous dissipation affecting flow focusing.
In this article, using experiments, direct numerical simulations, and scaling arguments, we characterize both the peak amplitude $F_1$ at impact and the one at take off ($F_2$) and elucidate their dependency on the control parameters: the Weber number $We$ (dimensionless impact velocity) and the Ohnesorge number $Oh$ (dimensionless viscosity). 
The first peak amplitude $F_1$ and the time $t_1$ to reach it depend on inertial timescales for low viscosity liquids, remaining nearly constant for viscosities up to 100 times that of water. For high viscosity liquids, we balance the rate of change in kinetic energy with viscous dissipation to obtain new scaling laws: $F_1/F_\rho \sim \sqrt{Oh}$ and $t_1/\tau_\rho \sim 1/\sqrt{Oh}$, where $F_\rho$ and $\tau_\rho$ are the inertial force and time scales, respectively, which are consistent with our data.
The time $t_2$ at which the amplitude $F_2$ appears is set by the inertio-capillary timescale $\tau_\gamma$, independent of both the viscosity and the impact velocity of the drop. However, these properties dictate the magnitude of this amplitude.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Schematic_v2.pdf}
	\caption{(a) Problem schematic with an axisymmetric computational domain used to study the impact of a drop with diameter $D_0$ and velocity $V_0$ on a non-wetting substrate. In the experiments, we use a quartz force sensor to measure the temporal variation of the impact force. The subscripts $d$ and $a$ denote the drop and air, respectively, to distinguish their material properties, which are the density $\rho$ and the dynamic viscosity $\eta$. The drop–air surface tension coefficient is $\gamma$. The grey dashed-dotted line represents the axis of symmetry, $r = 0$. Boundary air outflow is applied at the top and side boundaries (tangential stresses, normal velocity gradient, and ambient pressure are set to zero). The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). (b) The phase space with control parameters: the Weber number ($We$: dimensionless impact velocity) and the Ohnesorge number ($Oh$: dimensionless viscosity), exemplifying different applications.}
	\label{fig:schematic}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Figures/Figure1_summary_v5.pdf}
	\caption{Comparison of the drop impact force $F(t)$ obtained from experiments and simulations for the three typical cases with impact velocity $V_0 = 1.2\,\si{\meter}/\si{\second}, 0.97\,\si{\meter}/\si{\second}, 0.96\,\si{\meter}/\si{\second}$, diameter $D_0 = 2.05\,\si{\milli\meter}, 2.52\,\si{\milli\meter}, 2.54\,\si{\milli\meter}$, surface tension $\gamma = 72\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}, 61\,\si{\milli\newton}/\si{\meter}$ and viscosity $\eta_d = 1\,\si{\milli\pascal\second}, 25.3\,\si{\milli\pascal\second}, 80.2\,\si{\milli\pascal\second}$. These parameter give $Oh = 0.0025, 0.0625, 0.2$ and $We = 40$.
	For the three cases, the two peak amplitudes, $F_1/\rho_dV_0^2D_0^2 \approx$ 0.82, 0.92, 0.99 at $t_1 \approx 0.03\sqrt{\rho_dD_0^3/\gamma}$ and $F_2/\rho_dV_0^2D_0^2 \approx$ 0.37, 0.337, 0.1 at $t_2 \approx 0.42\sqrt{\rho_dD_0^3/\gamma}$, characterize the inertial shock from impact and the Worthington jet before takeoff, respectively. 
	The drop reaches the maximum spreading at $t_{\text{max}}$ when it momentarily stops and retracts until $t_3 \approx 0.8\sqrt{\rho_dD_0^3/\gamma}$ when the drop takes off ($F = 0$).
	Six instances are further elaborated through numerical simulations for ($We = 40, Oh = 0.0025$), namely (i) $t = 0\,\si{\milli\second}$ (touch-down), (ii) $t = 0.37\,\si{\milli\second}$ ($t_1$), (iii) $t = 2.5\,\si{\milli\second}$ ($t_{\text{max}}$), (iv) $t = 3.93\,\si{\milli\second}$ ($\approx 0.85t_2$), (v) $t = 4.63\,\si{\milli\second}$ ($t_2$), and (vi) $t = 5.25\,\si{\milli\second}$ ($\approx 1.15t_2$). 
	We stress the excellent agreement between experiments and simulations without any free parameters. The insets show representative snapshots at specific time instants overlaid with the drop boundaries from simulations in orange, also revealing good agreement. 
	The left part of each numerical snapshot shows the dimensionless local viscous dissipation function $\tilde{\xi}_\eta \equiv \xi_\eta D_0/\left(\rho_dV_0^3\right) = 2Oh\left(\boldsymbol{\tilde{\mathcal{D}}:\tilde{\mathcal{D}}}\right)$, where, $\boldsymbol{\mathcal{D}}$ is the symmetric part of the velocity gradient tensor, on a $\log_{10}$ scale and the right part the velocity field magnitude normalized with the impact velocity. The black velocity vectors are plotted in the center of mass reference frame of the drop to clearly elucidate the internal flow.}
\label{fig:summary}
\end{figure}

@@

Naturally, even the process of a Newtonian liquid drop impacting a rigid substrate is governed by a plethora of control parameters, including but not limited to the drop's density $\rho_d$, diameter $D_0$, velocity $V_0$, dynamic viscosity $\eta_d$, surface tension $\gamma$, and acceleration due to gravity $g$ (figure~\ref{fig:schematic}a). 
To navigate this rich landscape, we focus on two main dimensionless numbers that serve as control parameters (figure~\ref{fig:schematic}b): the Weber number $We$, which is the ratio of inertial to capillary forces and is given by

@@

\noindent Additionally, the Bond number

\begin{align}
	\label{Eq:BoDefinition}
	Bo = \frac{\rho_dgD_0^2}{\gamma}
\end{align}

\noindent compares gravity to inertial forces and is needed to uniquely define the non-dimensional problem.

@@

\begin{table}
	\begin{center}
		\def~{\hphantom{0}}
		\begin{tabular}{lccc}
			glycerol & $\rho_d$ & $\eta_d$  & $\gamma$\\
			(wt \%) &(kg/m$^{3}$) &(mPa.s)& (mN/m)\\[3pt]
			0 & 1000 & 1 & 72 \\
			50 & 1124 & 5 & 61 \\
			63 & 1158 & 10 & 61 \\
			74 & 1188 & 25.3 & 61 \\
			80 & 1200 & 45.4 & 61 \\
			85 & 1220 & 80.2 & 61 \\
		\end{tabular}
		\caption{Properties of the water-glycerol mixtures used in the experiments. $\rho_d$ and $\eta_d$ are the density and viscosity of the drop, respectively and $\gamma$ denotes the liquid-air surface tension coefficient. These properties are calculated using the protocol provided in \citet{cheng2008formula, volk2018density}.}
		\label{tab:table00}
	\end{center}
\end{table}
